---
title: What does your goodreads library say about you?
author: Adam Kirosingh
date: '2021-05-24'
slug: what-does-your-goodreads-library-say-about-you
categories:
  - Quantified Self
  - R
  - Python
tags:
  - Visualization
subtitle: ''
summary: ''
authors: []
lastmod: '2021-05-24T11:18:22-07:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
editor_options: 
  chunk_output_type: console
---
I've been recording the books that I've read in a csv since 2017 and decided to switch over to using the Goodreads website in 2019. While I don't find Goodreads particularly great at recommmending books, I do like the idea of using the data available to better understand reading habits. Before I would painstakingly add genre information and fill out information about the author and book for each entry, but goodreads provides a simple way to [export your data](https://help.goodreads.com/s/article/How-do-I-import-or-export-my-books-1553870934590).  

```{r include = F}
library(readr)
library(here)
library(tidyverse)
library(forcats)
library(viridis)
require(reshape2)
require(plyr)
require(scales)
library(xts)
```

In order to make the most of the data exported from Goodreads, Paul Klinger made a nice [python script](https://github.com/PaulKlinger/Enhance-GoodReads-Export) to scrap genre information from the website.

## Import the data 
```{r echo = T, warning=F,message=F}
Booklist <-
  read_csv(
    here::here(
      "content",
      "post",
      "2021-05-24-what-does-your-goodreads-library-say-about-you",
      "goodreads_library_export.csv"
    )
  )
```

## Tidy the data
```{r echo = T}
read_Booklist <- Booklist %>%
  subset(Booklist$`Exclusive Shelf` == "read") %>%
  mutate(genre_quorum = sub("\\|.*", "", genres))

short_Booklist <- read_Booklist %>%
  subset(read_Booklist$Author %in% names(sort(table(Booklist$Author), decreasing = T))[1:30])

#Turning Completion date into a date format R can recognize
read_Booklist$`Date Added`  <- as.Date(read_Booklist$`Date Added`,
                                       format = "%y-%m-%d")

read_Booklist$yearmonth <- as.yearmon(read_Booklist$`Date Added`)
read_Booklist$yearmonthf <- factor(read_Booklist$`Date Added`)
#Making a Week number column
read_Booklist$week <-
  as.numeric(strftime(read_Booklist$`Date Added`, format = "%V"))
# Making a month number column then creating a factor abbreviated month column. I found the factor step necessary to maintain order
read_Booklist$month <-
  as.numeric(strftime(read_Booklist$`Date Added`, format = "%m"))
read_Booklist$monthf <-
  factor(
    read_Booklist$month,
    levels = as.character(1:12),
    labels = c(
      "Jan",
      "Feb",
      "Mar",
      "Apr",
      "May",
      "Jun",
      "Jul",
      "Aug",
      "Sep",
      "Oct",
      "Nov",
      "Dec"
    ),
    ordered = TRUE
  )
#Repeating above for weekdays
read_Booklist$weekday <-
  strftime(read_Booklist$`Date Added`, format = "%w")
read_Booklist$weekdayf <-
  factor(
    read_Booklist$weekday,
    levels = rev(0:6),
    labels = rev(c(
      "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"
    )),
    ordered = TRUE
  )
read_Booklist$year <-
  strftime(read_Booklist$`Date Added`, format = "%Y")

summarized_Booklist <- read_Booklist %>%
  group_by(genre_quorum) %>%
  dplyr::summarise(Count = n()) %>%
  arrange(-Count)


calendar <- read_Booklist %>%
  mutate(monthweek = 1 + week - min(week)) %>%
  subset(year > 2017) %>%
  mutate(genre_quorum= factor(genre_quorum,levels = summarized_Booklist$genre_quorum))
```

## What genres do I tend to reads?

```{r genres, echo = T}
ggplot(data=read_Booklist, aes(x=fct_infreq(genre_quorum))) +
  geom_bar(stat="count",fill=c(viridis(33)))+ ylab("Number of Books")+ xlab("Goodreads Genre")+ ggtitle("Distribution of the genre of books I read") + theme_bw()  + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust =0.5))
```



## What's my favorite author?
```{r echo = T}
ggplot(data=short_Booklist, aes(x=fct_infreq(Author))) +
  geom_bar(stat="count",fill=c(viridis(29)))+ ylab("Number of Books")+ xlab("Author")+ ggtitle("Authors I've read multiple times") + theme_bw()  + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust =0.5))
```



## What do my readings look like?
```{r echo = T}
ggplot(calendar, aes(monthweek, weekdayf)) +
  geom_tile(colour = "white", aes(fill = genre_quorum)) +
  facet_grid(year ~ monthf, scales = "free") +
  labs(x = "Week of Month",
       y = "",
       title = "Finished Books Heatmap",
       subtitle = "Colored by book genre frequency",
       fill = "genre_quorum") + theme_bw() +  scale_fill_viridis_d() +theme(legend.position = "none", axis.text.x = element_blank(), axis.ticks.x = element_blank())
```


If you're interested in questions like, how does my ratings differ than the average or how many pages did I read last year, then try importing your csv into [Paul Klinger's website](https://almoturg.com/bookstats/). 
